\documentclass{article}
\begin{document}
How does one know if they are living in a democracy?  Is Russia a democracy?
Figuring out if you are living in an authoritarian state can involve more than just
what the constitution and laws say, but what is actually, going on.
Use this game as an indicator to see how democratic your country currently is.

Democracy Bingo Instructions
Every time the leader of your country does something that is only done in an authoritarian state,
you may be able to mark one of the bingo squares.  No action should cover more than one square.
If more than one criteria is met, choose the more specific rule.

Free Space is 14A Jan 6:
The 14th ammendment prohibits people that have engaged in insurrection from
holding public office.  Trump lead a violent insurrection on January 6, so he should be disqualified
from the presidency.  Note that the supreme court's decision in Trump v. Anderson only said that
States could not take Trump off the ballot, without saying anything about the insurrection,
so the Colorado finding holds here.
Parties guilty of ignoring the constitution:
Trump for running for Presidency
Supreme court for choosing to not state Trump's eligibility for presidency
Congress for not passing a law barring Trump from presidency.
Voters that voted for Trump, knowing the history and the ammendment.

14A other: If the President blatantly violates other parts of the 14th ammendment
then check this box.

22A Term Limit: If for any reason, Trump still has presidential power after Jan 20, 2029
then this is checked.

4A Unreasonable search and seisure:  If a policy isn't having the immediate positive outcome one
expects, then blame the pesky civil rights.  If searches without a search warrent become common, then
this may be checked.  This may include searches using drones or other high tech means.

5A Accused cannot be asked to be a witness against themselves
scenario: In order to make deportation more efficient, this right is removed from undocumented immigrants.

6A Fair trial:  Scenario: In order for Trump to obtain retribution against his political enemies, their right to trial
is removed.

8A Cruel, unusual or extreme punishment:
Scenario: In order for Trump to obtain retribution against his political enemies, extreme punishments are given
for minor infractions.  Looking hard enough, almost anyone can be found to have violated something.

10A Separation of federal and state powers:  If the president attempts to usurp states powers, then this is checked.

Other ammendments are listed as a number followed by an A.  Any blatant disregard for
a constitutional ammendment gets a box checked.

High crimes: Legitimate impeachment procedings are started, but do not affect the president.

Bribery:  President clearly and publicly offers or receives a bribe, but no action is
taken.

Overt Corruption: President is caught in a corrupt activity, that is not deniable, but
has no effect on office.  Typically an authoritarian ruler will say that the activity is
legal and perfectly acceptable.  May claim Presidential Immunity for all presidental
acts.
Example Scenario:  Elon Musk's super PAC spent approximately 200 Million helping Trump
get elected.
Trump may be creating a new position of ``Government Efficiency'' for Elon Musk.
Elon Musk runs companies that have been given fines, and has ongoing probes by FAA, EPA,
Labor Department, Justice Department, National Labor Relations board, Transportation Department
and others.  If the new Office of Government Efficiency were to tell all of these government
agencies, that to be efficient, they would need to drop all of these fines and investigations,
against Musk's companies, this would be a case of overt corruption.


FCC License: A standard authoritarian practice is to shut down all media that does not agree with the
head of state.  Trump has threatened to revoke the FCC license of major news channels.
If a major media outlet has its FCC license revoked, then mark this square.


Geneva Convention: (War Crimes): The president could instruct the military to commit war crimes.
Scenarios:  Israel has been accused of committing war crimes.  If the
US were to get actively involved, and Isreal is committing war crimes, then we could
transition from being an enabler of war crimes through our support of Israel to a
committer of war crimes.  If there is strong evidence of direct, active US participation
in violations of the Geneva Convention, then check this box.

Concentration Camps:  If a group of people are collected together, moved to a location
more convienient for the government, and intentionally subjected
to conditions significantly worse than jail, this may be considered
a concentration camp.
Scenario:  A poorly executed mass deportation in which people are collected
faster than they are taken out of the country.  This could lead to bad living conditions,
and a ``They got what they deserve'' attitude.  Consider the earlier process
of deporting a large group of people from where they were living in the US,
``Trail of Tears''.

Enemies: A very common method of maintaining power in authoritarian states, is to use the power
of the state to persecute your political enemies.
Criteria: 1 Innocent, because either there are no
actual accusations made, or there is no evidence supporting the accusation.
2  Expensive: the process must be costly for the victim, either in monetary terms,
reputation, or unjust imprisonment.  It can be very expensive
to mount a legal defense, and a person's personal or professional reputation may be ruined.
An example from the first Trump term is Hillary Clinton.
3  No remorse:  If someone is accused of a crime, but the investigation yields no supporting
evidence, then an apology, and compensation of some kind should be given.
If Trump intentionally unjustly persecutes political enemies, then this box can be checked.

Emergency: A classic method for illegitimately gaining power is to claim an emergency.
An attempt to gain power by claiming an emergency.
Scenarios: The president does not have the power to simply impose tarrifs.
Calling a situation an emergency could allow a president to exert undue power.
Claiming we are in an emergency in order to add a tarrif qualifies only if the tarrif is
actually added
for this reason.  There are a very wide range of cases covered by this trick.
Claiming a national emergency in order to not step down at the end of his term
would be considered Term-Limit

National Defense: Many things are justified as National Defense.
Scenarios: Getting tarriffs added to items that are domestically produced, and not militarily useful,
or necessary for survival can not be justified as national defense.  This is also a pretty broad
category

Missing Person:  In authoritarian states, the government will often kidnap or kill enemies, and
hide the crime.  If a political activist mysteriously dissappears, this may be checked.  Note
that this does not have to be done through official channels.  During Trumps trial, witnesses
received death threats, and Trump did nothing to try to stop them.  A sufficiently popular
president could inspire his supporters to take matters into their own hands.

Suspicious Death:  In authoritarian states, the government will often covertly kill political
enemies.  Exmples include suspicious plane crashes, poisonings, shootings, etc.

Revoke citizenship:  Another way for an authoritarian to make his problems go away is to revoke
the citizenship of a disliked group of people.

Military Suppression: Use of military as a response to peaceful protest.  Use of the 1807
insurrection act while legal, is a clear indication that there is a serious threat to
democracy involved.  If the military is used against protesters then this box can be checked.


Nepotism: In authoritarian states, rules place the welfare of their relatives over the
welfare of the state, and put them in powerful positions.  In 2020 Trump claimed that the
anti-nepotism law didn't apply to him.  If at least 4 people related Trump work in the white
house or are appointed to high level government positions because of Trump, then this
should be checked. 

Dogma: Formal policy requiring political or religious beliefs.  If it becomes a legal requirement
to agree with presidential policies or statements in order to get or hold a job, or receive
government services.  A religious test would also qualify for a check in this box.

Tip Line:
In order to keep the citizens of an authoritarian state in line, the leader often will
recruit its citizens to help with the task.  If an explicit ``Tip Line'' is set up allowing
individuals to report people that they know of being in violation of another
antidemocracy-bingo square, then this box can be checked.  In this age of artificial
intelligence, this does not need to be an explicit line to call.  If there is a known
system that monitors social media, and it is known that you can simply post your tip,
and it will be collected and used against the named person, then this still qualifies.
Scenarios:
A tip line where you can report that someone is against Trump or his policies,
in order to get him onto an Enemies List.
A line where you can report that someone does not believe something in order to get
him fired from his job.
A line where you can report that a project does not support the president's beliefs.
A mass deportation going on, that is inhumane (see Concentration camp), and
you can report that you believe someone is undocumented.
\end{document}
